Automatic Annotation of Tasks in Structured Code

### 2018-08-06 docs/ACM/paper.tex

\author{Fernando Magno Quint\~{a}o Pereira}
\authornote{This work is sponsored by CNPq, CAPEs, FAPEMIG, FAPESP, LG Electronics and
the CONTINUUM ANR Project.}          %% \authornote is optional;
%\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Professor}
%  \department{DCC}
  \institution{UFMG, Brazil \& LIRMM, France}
%  \streetaddress{6627 Antonio Carlos Avenue}
%  \city{Belo Horizonte}
%  \state{Minas Gerais}
%  \postcode{31.270-213}
%  \country{Brazil}
}
\email{fernando@dcc.ufmg.br}          %% \email is recommended

@@

Annotation systems have risen to a place of prominence as a simple and
effective means to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11} and
OmpSs~\cite{bueno:icpp:2011, duran:ppl:2011}.
Annotations let developers grant parallel
semantics to syntax originally written to execute sequentially.
%\guido{What do you mean by modern accelerators? Not clear.}
Combined with hardware accelerators such as GPUs and FPGAs, they have led
to substantial performance gains~\cite{Bertolli14,Mendonca17,Poesia17}.
%\guido{Suppport what?}
Nevertheless, although convenient, the use of annotations is not straightforward.
We still lack tools that help programmers to check if annotations
are correct and/or effective.
%replaced:
%Yet, even if convenient,

@@

This paper has described a methodology to annotate programs with task
parallel pragmas, which we could demonstrate to be effective
on general programs.
This methodology does not introduce any fundamentally new static analysis or code 
optimization; in this field, we claim no contribution.
Instead, our contributions lay into the overall design that eventually
emerged from a two-years long effort to combine existing compilation techniques
towards the goal of profiting from the powerful runtime system that OpenMP's
task parallelism brings forward.
During this period, we went through many dead-ends; too many, indeed, for a
11-pages report.
Nevertheless, this experience lets us single out a few elements from the
compiler literature, such as symbolic range analysis and windmills, which we
could use to solve challenges related to automatic code annotation.
Much work is still left to be done, until we can reach a stage in which
automatic annotations can beat consistently manual interventions.
In particular, our methodology asks for more aggressive tuning strategies.
Techniques such as Trancoso's~\cite{Trancoso17}, Iwasaki's~\cite{Iwasaki16B}
or Emani's~\cite{Emani15} have already been shown to be effective in this
domain, and we want to explore them further.
More specifically, a future release of \Taskminer{} is already expected to
feature Iwasaki's cut-off~\cite{Iwasaki16B}. \\
\textbf{Software} available at \url{http://cuda.dcc.ufmg.br/taskminer/}

>>>>

\author{Fernando Magno Quint\~{a}o Pereira}
\authornote{This work is sponsored by CNPq, CAPEs, FAPEMIG, FAPESP, ANR and LG Electronics.
While working on this project, Fernando was supported by a grant from the CONTINUUM ANR
Project (ANR-15-CE25-0007-0).}          %% \authornote is optional;
%\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Professor}
%  \department{DCC}
  \institution{UFMG, Brazil}
%  \streetaddress{6627 Antonio Carlos Avenue}
%  \city{Belo Horizonte}
%  \state{Minas Gerais}
%  \postcode{31.270-213}
%  \country{Brazil}
}
\email{fernando@dcc.ufmg.br}          %% \email is recommended

@@

Annotation systems have risen to a place of prominence as a simple and
effective way to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11} and
OmpSs~\cite{bueno:icpp:2011, duran:ppl:2011}.
Annotations let developers grant parallel
semantics to syntax originally written to execute sequentially.
%\guido{What do you mean by modern accelerators? Not clear.}
Combined with hardware accelerators such as GPUs and FPGAs, they have led
to substantial performance gains~\cite{Bertolli14,Mendonca17,Poesia17}.
%\guido{Suppport what?}
Nevertheless, although convenient, the use of annotations is not straightforward.
We still lack tools that help programmers to check if annotations
are correct and/or effective.
%replaced:
%Yet, even if convenient,

@@

This paper has described a methodology to annotate programs with task
parallel pragmas, which we could demonstrate to be effective
on general programs.
This methodology does not introduce any fundamentally new static analysis or code 
optimization; in this field, we claim no contribution.
Instead, our contributions lay into the overall design that eventually
emerged from a two-years long effort to combine existing compilation techniques
towards the goal of profiting from the powerful runtime system that OpenMP's
task parallelism brings forward.
During this period, we went through many dead-ends; too many, indeed, for a
11-pages report.
Nevertheless, this experience lets us single out a few elements from the
compiler literature, such as symbolic range analysis and windmills, which we
could use to solve challenges related to automatic code annotation.
Much work is still left to be done, until we can reach a stage in which
automatic annotations can beat consistently manual interventions.
In particular, our methodology asks for more aggressive tuning strategies.
Techniques such as Trancoso's~\cite{Trancoso17}, Iwasaki's~\cite{Iwasaki16B}
or Emani's~\cite{Emani15} have already been shown to be effective in this
domain, and we want to explore them further. \\

\noindent
\textbf{Software}:
\Taskminer{} can be used directly through an online interface available at
\url{http://cuda.dcc.ufmg.br/taskminer/}
